Title

### 2024-06-25 paper/main.tex
Updates

In this section, we develop a novel (conditional) measure of association for
mixed data. The presented measure extends partial correlation to mixed data by
combining a mixed-data residualization method with a multivariate measure of
association.

Given two variables $ X $ and $ Y $, and a set of conditional variables $
\bm{Z} $, we are interested in computing the conditional association, $ \rho(X,
Y; Z) $. As our measure is based on residulaization, we start by computing the
residuals for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor variables.
We use the mixed data residualization approach from \citet{Ankan2023}. We
consider each variable type separately and repeat the following steps for both
$ X $ and $ Y $:

\begin{enumerate}
	\item $ X $ is continuous: We start by training a regression model, $
		E_X: X \sim \bm{Z} $. Next we take the predictions from this
		model $ \hat{x} = E_X(\bm{Z}) $ and lastly the residuals are
		computed as the difference between the true and the predicted
		values.
		$$ R_X = X - \hat{x} $$
	\item $ X $ is ordinal: We start by training a probability estimator, $
		p_X: X \sim \bm{Z} $, and take probability predictions from it
		$ \hat{p}_X: p_X(\bm{Z}) $. Then we compute the residuals as follows:
		$$ R_X = $$
	\item $ X $ is categorical: We again start by training a probability estimator
		$ p_X: X \sim \bm{Z} $, and take the probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $.
		Next, we one-hot encode the categorical variable and then compute the residuals
		as follows:
		$$ R_X $$
\end{enumerate}

Here, users can choose the type of estimator that they want to use. There are
single estimators such as Random Forest or XGBoost that can work for both the
probability estimates and regression estimates. Or users can also choose
different estimators such as variants of linear regression for computing the
residuals. After the above steps are repeated for both $ X $ and $ Y $, we end
up with two residual matrices: $ R_X $ and $ R_Y $. The type of the variable
determines the shape for these matrices, i.e., if the variable is continuous we
get a residual matrix of shape $ ( n \times 1 ) $ and if the variable is
categorical we have a residual matrix of shape $ ( n \times (k-1)) $ (where one
of the dummy encoded variables is dropped to avoid multicollinearity)

>>>>

In this section, we develop a novel (conditional) measure of association for
mixed data by extending partial correlation to mixed data. We do this by
combining a mixed-data residualization method with a multivariate measure of
association. Given two variables $ X $ and $ Y $, and a set of conditional
variables $ \bm{Z} $, we are interested in computing an estimate of the
conditional association, $ \rho(X, Y; Z) $. Firstly, we compute the residuals
for $ x $ and $ y $ using the $ \bm{z} $ as the predictor variables. We use the
mixed data residualization approach from \citet{Ankan2023}. We consider each
variable type separately and repeat the following steps for both variables $ X
$ and $ Y $:

\begin{enumerate}
	\item \textbf{$ X $ is continuous:} We start by training a regression
		model, $ E_X: x \sim \bm{z} $. Next we take the predictions
		from this model $ \hat{x} = E_X(\bm{z}) $ and lastly the
		residuals are computed as the difference between the true and
		the predicted values.
		$$ R_{x_i} = x_i - \hat{x}_i $$
	\item \textbf{$ X $ is ordinal:} We start by training a probability
		estimator, $ p_X: x \sim \bm{z} $, and take probability
		predictions from it $ \hat{p}_X: \hat{p}_X(\bm{Z}) $. Then we
		compute the residuals as follows:
		$$ R_{x_i} = \hat{p}(\hat{x}_i > x_i) - \hat{p}(\hat{x}_i < x_i) $$
	\item \textbf{$ X $ is categorical:} We again start by training a
		probability estimator $ p_X: X \sim \bm{Z} $, and take the
		probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $. Next,
		we one-hot encode the categorical variable and then compute the
		residuals as follows: 
		$$ R_X = x_i - \hat{p}(\bm{z}_i) $$
\end{enumerate}

\todo[inline]{Fix notations above}

Here, users have the option to choose the estimators based on the properties of
their data. Non-parametric ensemble estimators such as Random Forest and
XGBoost can work well on a large variety of data and any data type. For linear
datasets users can also choose variants of linear regression depending on the
type of data. After the above steps are repeated for both $ X $ and $ Y $, we
end up with two residual matrices: $ R_X $ and $ R_Y $. The type of the
variable determines the shape for these matrices, i.e., if the variable is
continuous we get a residual matrix of shape $ ( n \times 1 ) $ and if the
variable is categorical we have a residual matrix of shape $ ( n \times (k-1))
$ (where one of the dummy encoded variables is dropped to avoid
multicollinearity)
